---
title: "Should We Be Going to Schools with Bad Food?"
author: "Chase Williamson and Tiffany Wang"
date: "Spring 2018"
output: html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Mini-Project 2: Data Wrangling

We're sure many of you have heard of Malcolm Gladwell, a famous pop social scientist who has written multiple books and now hosts a podcast called Revisionist History. In its first season, Gladwell aired an episode named [Food Fight](http://revisionisthistory.com/episodes/05-food-fight), highlighting the relationship between college endowment, money spent on food services, and number of low-income students at higher education instutions by comparing Vassar to Bowdoin. He argues that while Vassar cannot offer the luxury goods like hiring a chef to design delicious meals, its commitment to admitting low income students and making their education affordable is much more important and worth the sacrifice in terms of food quality.

While the contrast between Vassar and Bowdoin certainly provides evidence for a compelling narrative, it is important to look at the data to see if this is a trend across all institutions. Surely, it is possible to succeed on both fronts. We took a list of the [50 colleges](https://www.bestvalueschools.com/rankings/college-dining-halls/) with the best food in 2017 to compare them with various school stats from the U.S. Department of Education's [College Scorecard](https://collegescorecard.ed.gov/data/documentation/) data to verify Gladwell's claim.

#Grabbing the Data

We first scrape the list of 50 colleges with the best food in the country from the Best Value Schools website using the rvest package. 
```{r}
#Load packages
library(dplyr)
library(httr)
library(rvest)
library(tidyr)


#Assign url containing college food rankings to url.food
url.food <- "https://www.bestvalueschools.com/rankings/college-dining-halls/"

#Scrape food rankings from url
good.foods <- url.food %>%
  read_html() %>%
  html_nodes(css = "h3") %>%
  html_text()
```

Now, we have a list containing all text that was labeled as heading 3 on the webpage. We need to manipulate this dataset in two ways for it to be tidy: we must delete rows that are not true observations and we need to create one column for the name of the institution and another for the rank of the institution. We can delete the only row that does not correspond to an observation directly using the filter command from the dplyr package, and then we can use the separate command the tidyr package to create our two columns containing rank and institution name. 

```{r}
#Convert text object that we scraped to a dataframe with two columns: rank and school name
good.foods <- as.data.frame(good.foods) %>%
  filter(good.foods != "The 50 Colleges with the Best Food Ranking Methodology") %>%
  separate(col = good.foods, 
           into = c("food_rank", "school"), 
           extra = "merge")
```


We then scrape data from the Department of Education's college scorecard website using the rscorecard package. The college scorecard database is massive, and so we have to specify the specific variables that we want to select as well as the year. The most recent year for the college scorecard is 2015; we are not concerned about food quality changing between 2015 and 2017 because food quality is likely heavily dependent on facilities and expenditure which are unlikely to change significantly over a short period of time. The key variables of interes for us are percent of students who receive a Pell Grant and median student debt. 

```{r}
#Downloading the package to help load the data.
library(rscorecard)

#We have to use a key so that the government knows who's using their data
my_key <- "1HADPASnygYXC2p1ZxFMAN7j2VnS9IFDdvDMKZTW"

#Scrape data from college scorecard website
url <- "https://api.data.gov/ed/collegescorecard/v1/schools?"
scrape_scorecard <- GET(url, query = list(api_key = my_key))
sc_key(my_key)
dat <- sc_init() %>%
  sc_select(control, pctpell, instnm, DEBT_MDN) %>%
  sc_year(2015) %>%
  sc_get()
```


We then want to merge our food rankings dataset with the data from the college scorecard website. We also only care about whether a school is ranked in high quality food or not rather than the numerical rank of the school. Therefore we need to create a new indicator variable for whether the school was on the Best Value Schools list or not.

```{r}
#Join college scoreboard and food rankings dataset
colleges_food <- left_join(dat, good.foods, by = c("instnm" = "school"))

#Change food ranking into a numerical variable and create an indicator variable for if the school is ranked
colleges_food <- colleges_food %>%
  mutate(food_rank = as.numeric(food_rank), 
  food_ranked = as.numeric(!is.na(food_rank)))
```

#Let's Make Some Plots!

Now let's compare the percent of students receiving Pell Grants and students' median debt based on whether or not the school is ranked. To do this, we need the ggplot package to create a scatterplot.

```{r}
#Downloading ggplot 
library(ggplot2)

#Create a scatterplot of median debt against percent of students receiving pell grants
ggplot(colleges_food, aes(x = pctpell, y = log(debt_mdn), color = as.factor(food_ranked))) + 
  geom_point(alpha = .6) +
  labs(x = "Percent of Students Receiving Pell Grants",
       y = "Log of Median Debt",
       color = "Food Ranked?",
       title = "Does Food Quality Add Financial Burden?") +
  scale_color_manual(labels = c("No", "Yes"), 
                     values = c("#FF9A4A", "#108127")) +
  theme_light()
```

Most of the schools with food on the Best Value Schools list have points that lie on the upper left hand of the plot, indicating that there are less students receiving Pell Grants and that they have a higher median debt. This seems to fall in line with Gladwell's claim that schools that invest more in food also invest less in offering a place to low-income students. However, there are so many more non-ranked schools that it would make sense that there's a much larger spread of debt and student receiving Pell Grants.

#Comparing Similar Schools

One critique that might arise when looking at the relationship between the generosity of a school's financial aid package and food quality is that schools with really good food are fundamentally "different" from schools with lower quality food. We address this possible line of criticism by looking at the top 25 ranked schools under the notion that these will be similar institutions. We gathered information from [Forbes'](https://en.wikipedia.org/wiki/Forbes_Magazine%27s_List_of_America%27s_Best_Colleges) list of "America's Best Colleges" in 2017 and then merged this new dataset to our previously existing one.

```{r}
#Assign url containing college rankings to url.rank
url.rank <- "https://en.wikipedia.org/wiki/Forbes_Magazine%27s_List_of_America%27s_Best_Colleges"

#Scrape Forbes college rankings from wikipedia
forbes.top25 <- url.rank %>%
  read_html() %>%
  html_nodes(css = "li") %>%
  html_text() 

#Remove extraneous material
forbes.top25 <- as.data.frame(forbes.top25[c(1:25)])

#Create column containing college ranks. Note: that the colleges are in ranked order, but there is not a rank variable
forbes.top25$forbes_rank <- seq.int(nrow(forbes.top25))

#Label columns
colnames(forbes.top25) <- c("school", "forbes_rank")

#Merge college scorecard + food ranking data with college ranking data
colleges_food_rank <- left_join(colleges_food, 
                                forbes.top25, 
                                by = c("instnm" = "school"))

#Create "forbes_ranked" dummy variable indicating if the variable is labeled in that particular category
colleges_food_rank <- colleges_food_rank %>%
  mutate(forbes_rank = as.numeric(forbes_rank), 
         forbes_ranked = as.numeric(!is.na(forbes_rank)))

```

We created a similar plot to the one we had before, but now with only the schools in Forbes' "Best Colleges" list. With less data points, it's easier to see where all the points lie.

```{r}
#Look at the subset of colleges on the forbes list
colleges_food_rank_forbes <- colleges_food_rank %>%
  filter(forbes_ranked == 1)

#Create a scatterplot of median debt against percent of students receiving pell grants
ggplot(colleges_food_rank_forbes, aes(x = pctpell, y = log(debt_mdn), color = as.factor(food_ranked))) + 
  geom_point(alpha = .6) +
    labs(x = "Percent of Students Receiving Pell Grants",
       y = "Log of Median Debt",
       color = "Food Ranked?",
       title = "Forbes Top 25 Schools") +
  scale_color_manual(labels = c("No", "Yes"), 
                     values = c("#FF9A4A", "#108127")) +
  theme_light()
```

This plot is much less convincing than the first one. The schools with lowest median debt for students are all schools with ranked food, and there is a smaller discrepancy when looking at percent of students receiving Pell Grants. It seems like with the best schools (at least according to Forbes), quality of food does not affect your financial aid.

