---
title: "Should We Be Going to Schools with Bad Food?"
author: "Chase Williamson and Tiffany Wang"
date: "Spring 2018"
output: html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mini-Project 2: Data Wrangling

We're sure many of you have heard of Malcolm Gladwell, a famous pop social scientist who has written multiple books and now hosts a podcast called Revisionist History. In its first season, Gladwell aired an episode named [Food Fight](http://revisionisthistory.com/episodes/05-food-fight), highlighting the relationship between college endowment, money spent on food services, and number of low-income students at higher education instutions by comparing Vassar to Bowdoin. He argues that while Vassar cannot offer the luxury goods like hiring a chef to design delicious meals, its commitment to admitting low income students and making their education affordable is much more important and worth the sacrifice in terms of food quality.

While the contrast between Vassar and Bowdoin certainly provides evidence for a compelling narrative, it is important to look at the data to see if this is a trend across all institutions. Surely, it is possible to succeed on both fronts. We took a list of the [50 colleges](https://www.bestvalueschools.com/rankings/college-dining-halls/) with the best food in 2017 to compare them with various school stats from the U.S. Department of Education's [College Scorecard](https://collegescorecard.ed.gov/data/documentation/) data to verify Gladwell's claim.

We first scrape the list of 50 colleges with the best food in the country from the Best Value Schools website using the rvest package. 
```{r}
#Load packages
library(dplyr)
library(httr)
library(rvest)
library(tidyr)


#Assign url containing college food rankings to url.food
url.food <- "https://www.bestvalueschools.com/rankings/college-dining-halls/"

#Scrape food rankings from url
good.foods <- url.food %>%
  read_html() %>%
  html_nodes(css = "h3") %>%
  html_text()
```

Now, we have a list containing all text that was labeled as heading 3 on the webpage. We need to manipulate this dataset in two ways for it to be tidy: we must delete rows that are not true observations and we need to create one column for the name of the institution and another for the rank of the institution. We can delete the only row that does not correspond to an observation directly using the filter command from the dplyr package, and then we can use the separate command the tidyr package to create our two columns containing rank and institution name. 

```{r}
#Convert text object that we scraped to a dataframe with two columns: rank and school name
good.foods <- as.data.frame(good.foods) %>%
  filter(good.foods != "The 50 Colleges with the Best Food Ranking Methodology") %>%
  separate(col = good.foods, 
           into = c("food_rank", "school"), 
           extra = "merge")
```



We then scrape data from the Department of Education's college scorecard website using the rscorecard package. The college scorecard database is massive, and so we have to specify the specific variables that we want to select as well as the year. The most recent year for the college scorecard is 2015; we are not concerned about food quality changing between 2015 and 2017 because food quality is likely heavily dependent on facilities and expenditure which are unlikely to change significantly over a short period of time. The key variables of interes for us are type of institution (public, private not-for-profit, or private for-profit), percent of students who receive a Pell Grant, percent of first generation students, and cost of institution (there are separate variables for public and private universities). 

```{r}
#Downloading the package to help load the data.
library(rscorecard)

#We have to use a key so that the government knows who's using their data
my_key <- "1HADPASnygYXC2p1ZxFMAN7j2VnS9IFDdvDMKZTW"

#Scrape data from college scorecard website
url <- "https://api.data.gov/ed/collegescorecard/v1/schools?"
scrape_scorecard <- GET(url, query = list(api_key = my_key))
sc_key(my_key)
dat <- sc_init() %>%
<<<<<<< HEAD
  sc_select(control, pctpell, PAR_ED_PCT_1STGEN, instnm, npt4_pub, npt4_priv, DEBT_MDN) %>%
=======
  sc_select(control, pctpell, PAR_ED_PCT_1STGEN, instnm, npt4_pub, npt4_priv, GRAD_DEBT_MDN) %>%
>>>>>>> bc1339b37c6f1650c69d226de90bd7b78755c11c
  sc_year(2015) %>%
  sc_get()

#Combining the two variables of cost of public and private institutions 
dat <-  gather(dat, npt4_pub, npt4_priv, key = "Type of School", value = "Cost")
```


We then want to merge our food rankings dataset with the data from the college scorecard website.
```{r}
colleges_food <- left_join(dat, good.foods, by = c("instnm" = "school"))
```

One critique that might arise when looking at the relationship between the generosity of a school's financial aid package and food quality is that schools with really good food are fundamentally "different"from schools with lower quality food. We address this possible line of criticism by looking at the top 25 ranked schools under the notion that these will be similar institutions. We gathered information from [Forbes'](https://en.wikipedia.org/wiki/Forbes_Magazine%27s_List_of_America%27s_Best_Colleges) list of "America's Best Colleges" in 2017 and then merged this new dataset to our previously existing one.
```{r}
#Assign url containing college rankings to url.rank
url.rank <- "https://en.wikipedia.org/wiki/Forbes_Magazine%27s_List_of_America%27s_Best_Colleges"

#Scrape Forbes college rankings from wikipedia
forbes.top25 <- url.rank %>%
  read_html() %>%
  html_nodes(css = "li") %>%
  html_text() 

#Remove extraneous material
forbes.top25 <- as.data.frame(forbes.top25[c(1:25)])

#Create column containing college ranks. Note: that the colleges are in ranked order, but there is not a rank variable
forbes.top25$forbes_rank <- seq.int(nrow(forbes.top25))

#Label columns
colnames(forbes.top25) <- c("school", "forbes_rank")

#Merge college scorecard + food ranking data with college ranking data
colleges_food_rank <- left_join(colleges_food, 
                                forbes.top25, 
                                by = c("instnm" = "school"))

#Create "food_ranked" and "forbes_ranked" dummy variable indicating if the variable is labeled in that particular category
colleges_food_rank <- colleges_food_rank %>%
  mutate(food_rank = as.numeric(food_rank), 
         forbes_rank = as.numeric(forbes_rank)) %>%
  mutate(food_ranked = as.numeric(!is.na(food_rank)), 
         forbes_ranked = as.numeric(!is.na(forbes_rank))) 

```

In this R chunk, we look at a few different visualizations
```{r}
#Downloading ggplot to make plots.
library(ggplot2)
ggplot(colleges_food_rank, aes(x = pctpell, y = log(debt_mdn), color = as.factor(food_ranked))) + 
  geom_point(alpha = .6)

#What if we just look at school categories?
private_food <- colleges_food_rank %>%
  filter(!is.na(pctpell), !is.na(grad_debt_mdn)) %>%
  group_by(control, food_ranked) %>%
  summarise(mean_pctpell = mean(pctpell), 
            mean_grad_debt_mdn = mean(grad_debt_mdn))

ggplot(private_food, aes(x = mean_grad_debt_mdn, y = mean_pctpell, color = as.factor(food_ranked))) + 
  geom_point()

#What if we only look at schools that are ranked in the top 25 by forbes? 
forbes_ranked_only <- filter(colleges_food_rank, forbes_ranked==1)
ggplot(forbes_ranked_only, aes(x = par_ed_pct_1stgen, y = pctpell, color = as.factor(food_ranked))) + 
  geom_point()

#Does the cost of the school have to do with the quality of food?
ggplot(colleges_food_rank, aes(x = as.character(food_ranked), y = Cost)) + 
  geom_boxplot()

ggplot(colleges_food_rank, aes(x = as.character(food_ranked), y = grad_debt_mdn)) + 
  geom_boxplot()

#But how about when we just take the "good" schools?
ggplot(forbes_ranked_only, aes(x = as.character(food_ranked), y = Cost)) + 
  geom_boxplot()

<<<<<<< HEAD
#But how about when we just take the "good" schools?
ggplot(forbes_ranked_only, aes(x = as.character(food_ranked), y = debt_mdn)) + 
=======
ggplot(forbes_ranked_only, aes(x = as.character(food_ranked), y = grad_debt_mdn)) + 
>>>>>>> bc1339b37c6f1650c69d226de90bd7b78755c11c
  geom_boxplot()
```

In this R chunk, we show summary stats
```{r}



```


